Cosmological constraints from the redshift dependence of the galaxy angular 2-point correlation function

### 2016-10-06 draft/draft1.tex
updates

\begin{abstract}
We propose to use the redshift dependence of the galaxy angular 2-point correlation function (2pCF)
for a purpose of cosmological constraints.
When an incorrect cosmological model is assumed for the coordinate transformation from redshift space to comoving space, 
it produces {\it redshift-dependent mis-scaling} in the constructed galaxy distribution.
As a result, we find a redshift-dependent rescaling in the measured galaxy angular 2pCF .
We test our method on the HR4 mock galaxies having hundreds of millions of galaxies at redshifts 
$0\leq z\leq2$.
We find the redshift evolution of shape of angular 2pCF, as a function of scale,
is sensitive to the adopted cosmological parameters,
and is relatively insensitive to the gravitational growth of structure,
the galaxy bias, and the redshift space distortion.
Analyzing the 2pCF shape on scales $5  h^{-1} {\rm Mpc} \leq r_\perp \leq 40 h^{-1} {\rm Mpc}$, 
we derive tight constraints on $\Omega_m$ and $w$ with 68.3\% CL uncertainties of 0.015 and 0.05.
%Our method is complementary to current works using 2pCF to constrain cosmological parameters.
The method could be applicable to future large scale structure surveys, especially the photometric surveys such as DES, LSST, 
to derive tight cosmological constraints.
This work is a continuation of our previous works \citep{Li2014,Li2015,Li2016}
as a strategy to constraining cosmological parameters based on some nearly redshift-invariant physical quantity;
they are all powerful in constraining cosmology and have a lot of merit,
e.g., are less affected by the systematic effects introduced by non-linear clustering or RSD,
do not require accurate analytical modeling of the galaxy correlation function,
and are applicable on relatively small clustering scales.
\end{abstract}

@@

In \cite{Li2014,Li2015,Li2016} we developed a novel strategy to probe $D_A$ and $H$ from LSS data.
%proposed a series of methods to probe the geometric information encoded in the LSS data.
We found that the AP or volume effect in general introduces a geometric distortion that evolves significantly with redshift;
this phenomenon can be utilized to distinguish the AP effect from the other effects, such as the redshift space distortion (RSD).
The radial distances of galaxies are inferred from the measured redshifts, which are distorted by the peculiar motion of galaxies, 
leading to apparent distortion in the redshift space galaxy distribution \citep{FOG,Kaiser1987,Ballinger1996}.
This is the major systematic effect limiting our ability to probe the geometry of LSS from galaxy distribution.
We found its influence is significantly reduced when studying the redshift dependence of the distortion.

@@

In this paper we continue our exploration of this strategy and presented 
a new method using the redshift of the shape of the {\it galaxy angular 2pCF} as a function of radial scale, $\omega(r_{\bot})$,
to constrain cosmology.
A wrongly adopted cosmology result in {\it redshift-dependent mis-scaling} of the galaxy distribution,
which leads to redshift-evolution of the shape of $\omega(r_{\bot})$.
The signal is sensitive to the cosmological parameters and relatively insensitive to the other effects such as the gravitational growth of structure,
the galaxy bias, and the RSD effect.
We apply our method to HR4 mock galaxies, using hundreds of millions of galaxies at redshfits $0\leq z\leq2$
and develop the statistical procedure to constrain cosmological parameters.

@@

In order to directly compare the shape of the 2pCF at different redshifts,
in the middle panel of Figure \ref{fig_diffz} 
we show the $r_\perp\omega$ normalized by the overall amplitude within 
$5h^{-1}{\rm Mpc} < r < 40h^{-1}{\rm Mpc}$
(here after $\hat \omega_{r_\perp}$):
\begin{equation}
 \hat \omega_{ r_\perp} \equiv  \frac{ {r_\perp\omega}(r_{\perp})}{ \int_{r_{\perp, \rm min}}^{r_{\perp, \rm max}}r_{\perp} \omega(r_{\perp}) d r_{\perp} / (r_{\perp,\rm max}-r_{\perp,\rm min}) },
\end{equation}
where we choose $r_{\perp, \rm min}=5 h^{-1}$ Mpc, $r_{\perp, \rm max}=40 h^{-1}$ Mpc in this analysis.
Below $5 h^{-1}$ Mpc the non-linear growth of structure 
leads to systematic redshift evolution which could be difficult to be reliably accounted;
On scales larger than $40 h^{-1}$ Mpc,
the analytical modelling of the shape of the angular 2pCF is relatively well understood;
one can just fit the 2pCF with the theoretical predictions \citep{Salvador2014,Salvador2016} 
instead of using our method 
(although our method should also be applicable).

@@

\begin{figure*}
   \centering{
%   \includegraphics[width=18cm]{fig9_0.eps}
%   \includegraphics[width=18cm]{fig9_1.eps}
%   \includegraphics[height=8cm]{Tpcf--plot--Normed.eps}
%    \includegraphics[height=8cm]{smu.eps}
    \includegraphics[width=18cm]{fig5.eps}
   }
   \caption{\label{fig_cosmo}
    $\hat{\omega}_{r_\perp}$ (upper panels) and $\hat{\omega}_{r_\perp}(z=z_i) - \hat{\omega}_{r_\perp}(z=0)$ (lower panels) at several redshifts,
    for the four wrong cosmologies.
    The redshift evolution of angular 2pCF shape in these wrong cosmologlies are detected at high CL.
   }
\end{figure*}

@@

\begin{figure*}
   \centering{
   %\includegraphics[width=7cm]{Tpcf--contour.eps}
   %\includegraphics[width=16cm]{Tpcf--contour-others.eps}
   %\includegraphics[width=16cm,natwidth=52,natheight=40]{fig10.eps}
   \includegraphics[width=16cm]{fig9.eps}
   }
   \caption{\label{fig_contours}
   Likelihood contours (68.3\%, 95.4\%) in the $\Omega_m-w$ plane from our method,
   a likelihood value computed using Equations (\ref{eq:like,eq:chisq1}), 
   based on the redshfit evolution of the angular 2pCF shape $\hat\omega_{r_\perp}$.
   Utilizing the HR4 snapshots data of mock galaxies at five redshifts,
   the method can lead to very tight cosmological constraint.
   }
\end{figure*}

We obtain tight constraints on the two paramters.
The 2$\sigma$ contour lies within the region $0.23<\Omega_m<0.285$, $-1.1<w<-0.9$.
The two parameters are very strongly degenerated with each other.
The thin shape of the contour implies that, 
when combining with the another observational data with a different direction of degeneracy (e.g. CMB, BAO),
a very tight combined constraint can be obtained.
In the case of fixing one parameter at its best-fit value and constraining the other,
one obtains 1$\sigma$ uncertainties of $\delta\Omega_m\approx0.002,\delta w\approx0.01$.

>>>>

\begin{abstract}
We propose to use the redshift dependence of the galaxy angular 2-point correlation function (2pCF)
for a purpose of cosmological constraints.
When an incorrect cosmological model is assumed for the coordinate transformation from redshift space to comoving space, 
it produces {\it redshift-dependent mis-scaling} in the constructed galaxy distribution.
As a result, we observe a redshift-dependent rescaling of the measured galaxy angular 2pCF.
We test our method on the HR4 mock galaxies having hundreds of millions of galaxies at redshifts 
$0\leq z\leq2$.
We find the redshift evolution of shape of angular 2pCF, as a function of scale,
is sensitive to the adopted cosmological parameters,
and is relatively insensitive to the gravitational growth of structure,
the galaxy bias, and the redshift-space distortion.
Analyzing the 2pCF shape on scales $5  h^{-1} {\rm Mpc} \leq r_\perp \leq 40 h^{-1} {\rm Mpc}$, 
we derive tight constraints on $\Omega_m$ and $w$ with 68.3\% CL uncertainties of 0.015 and 0.05.
%Our method is complementary to current works using 2pCF to constrain cosmological parameters.
The method could be applicable to future large scale structure surveys, especially the photometric surveys such as DES, LSST, 
to derive tight cosmological constraints.
This work is a continuation of our previous works \citep{Li2014,Li2015,Li2016}
as a strategy to constraining cosmological parameters based on some nearly redshift-invariant physical quantity;
they are all powerful in constraining cosmology and have a lot of merit,
e.g., are less affected by the systematic effects introduced by non-linear clustering or RSD,
do not require accurate analytical modeling of the galaxy correlation function,
and are applicable on relatively small clustering scales.
\end{abstract}

@@

In \cite{Li2014,Li2015,Li2016} we developed a novel strategy to probe $D_A$ and $H$ from LSS data.
%proposed a series of methods to probe the geometric information encoded in the LSS data.
We found that the AP or volume effect in general introduces a geometric distortion that evolves significantly with redshift;
this phenomenon can be utilized to distinguish the AP effect from the other effects, such as the redshift-space distortion (RSD).
The radial distances of galaxies are inferred from the measured redshifts, which are distorted by the peculiar motion of galaxies, 
leading to apparent distortion in the redshift space galaxy distribution \citep{FOG,Kaiser1987,Ballinger1996}.
This is the major systematic effect limiting our ability to probe the geometry of LSS from galaxy distribution.
We found its influence is significantly reduced when studying the redshift dependence of the distortion.

@@

In this paper we continue our exploration of this strategy and presented 
a new method using the redshift of the shape of the {\it galaxy angular 2pCF} as a function of radial scale, $\omega(r_{\bot})$,
to constrain cosmology.
A wrongly adopted cosmology result in {\it redshift-dependent mis-scaling} of the galaxy distribution,
which leads to redshift-evolution of the shape of $\omega(r_{\bot})$.
The signal is sensitive to the cosmological parameters and relatively insensitive to the other effects such as the gravitational growth of structure,
the galaxy bias, and the RSD effect.
We apply our method to HR4 mock galaxies, using hundreds of millions of galaxies at redshifts $0\leq z\leq2$
and develop the statistical procedure to constrain cosmological parameters.

@@

In order to directly compare the shape of the 2pCF at different redshifts,
in the middle panel of Figure \ref{fig_diffz} 
we show the $r_\perp\omega$ normalized by the overall amplitude within 
$5h^{-1}{\rm Mpc} < r < 40h^{-1}{\rm Mpc}$
(here after $\hat \omega_{r_\perp}$):
\begin{equation}
 \hat \omega_{ r_\perp} \equiv  \frac{ {r_\perp\omega}(r_{\perp})}{ \int_{r_{\perp, \rm min}}^{r_{\perp, \rm max}}r_{\perp} \omega(r_{\perp}) d r_{\perp} / (r_{\perp,\rm max}-r_{\perp,\rm min}) },
\end{equation}
where we choose $r_{\perp, \rm min}=5 h^{-1}$ Mpc, $r_{\perp, \rm max}=40 h^{-1}$ Mpc in this analysis.
Below $5 h^{-1}$ Mpc the non-linear growth of structure 
leads to systematic redshift evolution which could be difficult to be reliably accounted;
On scales larger than $40 h^{-1}$ Mpc,
the analytical modeling of the shape of the angular 2pCF is relatively well understood;
one can just fit the 2pCF with the theoretical predictions \citep{Salvador2014,Salvador2016} 
instead of using our method 
(although our method should also be applicable).

@@

\begin{figure*}
   \centering{
%   \includegraphics[width=18cm]{fig9_0.eps}
%   \includegraphics[width=18cm]{fig9_1.eps}
%   \includegraphics[height=8cm]{Tpcf--plot--Normed.eps}
%    \includegraphics[height=8cm]{smu.eps}
    \includegraphics[width=18cm]{fig5.eps}
   }
   \caption{\label{fig_cosmo}
    $\hat{\omega}_{r_\perp}$ (upper panels) and $\hat{\omega}_{r_\perp}(z=z_i) - \hat{\omega}_{r_\perp}(z=0)$ (lower panels) at several redshifts,
    for the four wrong cosmologies.
    The redshift evolution of angular 2pCF shape in these wrong cosmologies are detected at high CL.
   }
\end{figure*}

@@

\begin{figure*}
   \centering{
   %\includegraphics[width=7cm]{Tpcf--contour.eps}
   %\includegraphics[width=16cm]{Tpcf--contour-others.eps}
   %\includegraphics[width=16cm,natwidth=52,natheight=40]{fig10.eps}
   \includegraphics[width=16cm]{fig9.eps}
   }
   \caption{\label{fig_contours}
   Likelihood contours (68.3\%, 95.4\%) in the $\Omega_m-w$ plane from our method, 
   based on the redshfit evolution of the angular 2pCF shape $\hat\omega_{r_\perp}$.
   Utilizing the HR4 snapshots data of mock galaxies at five redshifts,
   the method can lead to very tight cosmological constraint.
   }
\end{figure*}

We obtain tight constraints on the two parameters.
The 2$\sigma$ contour lies within the region $0.23<\Omega_m<0.285$, $-1.1<w<-0.9$.
The two parameters are very strongly degenerated with each other.
The thin shape of the contour implies that, 
when combining with the another observational data with a different direction of degeneracy (e.g. CMB, BAO),
a very tight combined constraint can be obtained.
In the case of fixing one parameter at its best-fit value and constraining the other,
one obtains 1$\sigma$ uncertainties of $\delta\Omega_m\approx0.002,\delta w\approx0.01$.
